Simulation and Optimization of Injector Parameters with OpenFOAM and Globalizer Tools

### 2021-08-03 2021/ICNAAM/2021_ICNAAM.tex
Corrections in Russian

\begin{abstract}
The paper presents the results of solving the problem of simulation and research of processes taking place in a fuel system injector. Simulation has been carried using the tools of the OpenFOAM system. The problem of differential pressure minimization at the injector inlet and outlet was set and solved. This problem was a Lipschitz optimization problem with a black-box type objective function. To solve it, we used an efficient global search algorithm implemented in the Globalizer system.
\end{abstract}

@@

As engineering objects and conditions of their functioning become more complicated, it results in a significant increase in the complexity of corresponding mathematical models and numerical methods for their analysis. Nowadays, the main (and often the only possible) tool for such analysis is supercomputer simulation of the object's behavior. For this purpose, CAD/CAM/CAE-systems are widely used. A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical simulation of continuum mechanics problems \cite{OpenFOAM}).

@@

The optimal parameter selection problems under consideration are characterized by the fact that the objective function $\varphi(y)$ is not defined analytically; there is only some algorithm to calculate its values at the points of the domain $D$.  In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Kalyulin2017}.

@@

A typical assumption used in many global optimization methods \cite{Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996} is the assumption that the objective function $\varphi(y)$ satisfies the Lipschitz condition
\[
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D, 0<L<\infty.
\]

@@

One approach to dimensionality reduction is to use Peano-type space filling curves (\textit{evolvents}). Using evolvents $y(x)$, which continuously and uniquely map a unit interval onto a hypercube, reduces the problem of finding the minimum of the function $\varphi(y)$ in the domain $D$ to finding the minimum of a one-dimensional function  on the interval $[0,1]$
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
\]

The above assumptions and approaches have led to the development of a number of global search algorithms based on the information-statistical approach \cite{Strongin2000}. These algorithms are implemented in the Globalizer software system for solving global optimization problems \cite{globalizerSystem}. Here, we present a scheme of the basic global search algorithm (GSA). This is a Divide-the-Best algorithm, i.e. it is aimed to conduct a subsequent search trial in the most promising (from the point of view of global minimum detection) search subdomain.

@@

Step 2. For each interval $(x_{i-1},x_i), \; 1\leq i \leq k,$ compute the characteristics $R(i)$ using some formulae.

Step 3. Find the interval $(x_{t-1},x_t),$ with the maximum characteristic 
\[
R(t)=\max\{R(i):1 \leq i \leq k+1\}.
\]

@@

@@

In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields. The injector simulation was divided into several key steps:
\begin{enumerate}
\item Mathematical formulation. This includes a procedure for selecting equations describing the dynamics of a fluid or gas, setting boundary conditions and initial conditions. It does not change from iteration to iteration.
\item Parameterization. Selection of varying parameters that will vary from iteration to iteration.
\item Generating the geometry and the computational grid. In line with Parameterization, a procedure for renewing the geometry and computational grid is performed.
\item Calculation of hydrodynamic fields. At each iteration, the system of equations describing fluid dynamics is solved numerically.
\item Interpretation. Physical interpretation of the results after the optimization procedure.
\end{enumerate}

@@

The geometric representation of the computational domain is shown in Figure 1, with the parameterization points highlighted in red. The geometry is parametrized by increasing the distance between the red points symmetrically with respect to the channel centreline.

@@

The calculation was carried out with the standard pisoFoam solver and the PISO algorithm \cite{Issa1986_2,Issa1986_1}.
A total of 300 finite elements were involved in the calculation. The process of finding the effective channel thickness took 3395 seconds on an Intel(R) Xeon(R) CPU X5670 2.93 GHz, during which time 69 iterations of the global search algorithm were performed. The horizontal component of the velocity at the entrance to the computational domain $\vec{U_i}$  was 10 m/s. The pressure at the exit of the computational domain $p_o$ was equal to zero, just like all components of the velocity vector at the exit $\vec{U_o}$ . Linear dimensions of the computational domain are shown in the schematic of Figure 1. As a result of solving the problem, the optimum channel thickness (0.59 cm) was found, giving the minimum fuel consumption.

>>>>

\begin{abstract}
The paper presents results of solving the problem of simulation and research of processes taking place in a fuel system injector. Simulation has been carried using tools of the OpenFOAM system. The problem of differential pressure minimization at the injector inlet and outlet was set and solved. This problem was considered as a Lipschitz optimization problem with a black-box type objective function. To solve it, we used an efficient global search algorithm implemented in the Globalizer system.
\end{abstract}

@@

As engineering objects and conditions of their functioning become more complicated, it results in a significant increase in the complexity of corresponding mathematical models and numerical methods for their analysis. Nowadays, the main (and often the only possible) tool for such analysis is supercomputer simulation of the object's behavior. For this purpose, CAD/CAM/CAE-systems are widely used. A well-known example of open source software of this kind is OpenFOAM, an open platform for numerical simulation of continuum mechanics problems \cite{OpenFOAM}.

@@

The optimal parameter selection problems under consideration are characterized by the fact that the objective function $\varphi(y)$ is not defined analytically; there is only some algorithm to calculate its values at the points of the domain $D$.  In this case, one search trial (calculating the function value in a particular point of the search domain) is a computationally intensive operation \cite{Sergeyev1999,Kvasov2013,Kalyulin2017,Sergeyev2018,Paulavicius2020,Lera2021}.

@@

A typical assumption used in many global optimization methods \cite{Sergeyev2013,Evtushenko2013,Jones2009,Zilinskas2010,Pinter1996,Grishagin2016_1,Sergeyev2021} is the assumption that the objective function $\varphi(y)$ satisfies the Lipschitz condition
\[
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D, 0<L<\infty.
\]
% Предположение такого рода является достаточно естественным для многих прикладных задач, поскольку относительные вариации целевой функции, характеризующей моделируемую систему, обычно не могут превышать некоторый порог, определяемый ограниченной энергией изменений в системе.

@@

One approach to dimensionality reduction is to use Peano-type space filling curves (\textit{evolvents}). Using evolvents $y(x)$, which continuously and uniquely map a unit interval onto a hypercube, reduces the problem of finding the minimum of the function $\varphi(y)$ in the domain $D$ to finding the minimum of a one-dimensional function  on the interval $[0,1]$
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}}.
\]
As follows from the properties of the space-filling curve \cite{Sergeyev2013}, the reduced one-dimensional function $\varphi(y(x))$ satisfies a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}, \; x_1,x_2 \in [0,1],
\]
where the H{\"o}lder constant $H$ is related to the Lipschitz constant $L$ by the relationship $H=2L\sqrt{N+3}$.

The above assumptions and approaches have led to the development of a number of global search algorithms based on the information-statistical approach \cite{Strongin2000}. These algorithms are implemented in the Globalizer software system for solving global optimization problems \cite{globalizerSystem}. Here, we present a scheme of the basic global search algorithm (GSA). This is a Divide-the-Best algorithm \cite{Sergeyev1998}, i.e. it is aimed to conduct a subsequent search trial in the most promising (from the point of view of global minimum detection) search subdomain.

@@

Step 2. For each interval $(x_{i-1},x_i), \; 1\leq i \leq k+1,$ compute the characteristics $R(i)$ using some formulae.

Step 3. Find the interval $(x_{t-1},x_t),$ with the maximum characteristic \cite{Grishagin1997}
\[
R(t)=\max\{R(i):1 \leq i \leq k+1\}.
\]

@@

The global search algorithm has the folloing interpretation. Using trial points it is possible to design a minorant for which characteristic $R(i)$ is the minimum value of the minorant in the interval $(x_{i-1}, x_i ), 1 \leq i \leq k + 1,$ taken with the opposite sign. Then new trial is carried out in the interval where the global minimizer of the current minorant is situated and the new trial point $x^{k+1}$ coincides with it.

@@

In our particular case, each iteration of the optimal solution search process was a separate numerical calculation of the hydrodynamic fields. The injector simulation was divided into several key steps:
\begin{enumerate}
\item \textbf{Mathematical formulation}. This includes a procedure for selecting equations describing the dynamics of a fluid or gas, setting boundary conditions and initial conditions. It does not change from iteration to iteration.
\item \textbf{Parameterization}. Selection of varying parameters that will vary from iteration to iteration.
\item \textbf{Generating the geometry and the computational grid}. In line with Parameterization, a procedure for renewing the geometry and computational grid is performed.
\item \textbf{Calculation of hydrodynamic fields}. At each iteration, the system of equations describing fluid dynamics is solved numerically.
\item \textbf{Interpretation}. Physical interpretation of the results after the optimization procedure.
\end{enumerate}

@@

The geometric representation of the computational domain is shown in Figure 1, with the parameterization points highlighted in red. The geometry is parametrized by increasing the distance between the red points symmetrically with respect to the channel centreline. 
%Таким образом, в задаче имеется один варьируемый параметр (диаметр сечения канала в его средней части), значения которого изменяются в диапазоне $[0,10]$.

@@

The calculation was carried out with the standard pisoFoam solver and the PISO algorithm \cite{Issa1986_2,Issa1986_1}.
A total of 300 finite elements were involved in the calculation. The process of finding the effective channel thickness took 3395 seconds on an Intel(R) Xeon(R) CPU X5670 2.93 GHz, during which time 69 iterations of the global search algorithm were performed. 
% Для решения задачи был использован смешанный локально-глобальный алгоритм \cite{Barkalov2010}, реализованный в Globalizer.
% При этом в критерии остановки метода было использовано значение $\epsilon = 10^{-3}$.
The horizontal component of the velocity at the entrance to the computational domain $\vec{U_i}$  was 10 m/s. The pressure at the exit of the computational domain $p_o$ was equal to zero, just like all components of the velocity vector at the exit $\vec{U_o}$ . Linear dimensions of the computational domain are shown in the schematic of Figure 1. As a result of solving the problem, the optimum channel thickness (0.59 cm) was found, giving the minimum fuel consumption.

%Найденное решение является удовлетворительным с точки зрения решаемой задачи. Например, другие решения, которые получаются при запуске методов локальной оптимизации для решения данной задачи, зависят от выбора стартовой точки. И значение целевой функции при этом получается не меньше, чем значение в точке, найденной Globalizer.
